\section*{Author’s roles}
%Manuscripts must include details for the contributions of each of the authors, including participation in study design, execution, analysis, manuscript drafting and critical discussion (refer to ‘Authorship’ section).
Q.A., A.C., S.D.B., and V.C. conceived and designed the study. 
S.B., A.D.M., G.D.M., M.P., and V.C. wrote the code and performed the bioinformatics analyses.
M.C. and C.F. provided support to bioinformatics analyses. 
V.A., A.R., I.D.B., P.D.A., and G.E. performed the experiments.  
P.G., and M.R. contributed to the clinical samples and collected clinical data. 
S.B. and V.C. wrote the manuscript. 
All authors critically reviewed the manuscript and approved the final version.

\section*{Acknowledgements}
%Personal acknowledgements should precede those of institutions or agencies.
We are thankful to all volunteers that were enrolled in the study, as well as all the medical personnel that contributed to the collection of samples. We are also thankful to Prof. Nicole Soranzo and Prof. Erik Garrison for help in the early stage of the project.   


\section*{Funding}
%With respect to funding of research, in line with the World Association of Medical Editors (WAME) guidelines, the journal considers it the responsibility of the author to protect the integrity of the research record from bias related to the source of funding by fully declaring all sponsorships, the roles played by sponsors in the research as well as institutional affiliations and relevant financial ties.
%These should be listed in the manuscript in the Funding section. Grant numbers should be given.
%Where no specific funding was sought for the study, and departmental funds were used to support the authors throughout the study period and manuscript preparation, this should be acknowledged, giving department names.
P.O.R. Campania FSE 2014-2020 and EMBO STF 7919 to V.C. The computational work has been executed on the IT resources of the ReCaS-Bari data center, which have been made available by two projects financed by the MIUR (Italian Ministry for Education, University and Re-search) in the \textit{PON Ricerca e Competitività 2007-2013" Program: ReCaS (Azione I - Interventi di rafforzamento strutturale, PONa3\_00052, Avviso 254/Ric) and PRISMA (Asse II - Sostegno all'innovazione, PON04a2\_A)}

\section*{Conflict of interest}
%Authors should include a conflict of interest statement on the manuscript detailing any potential conflicts of interest of any of the authors due to relationships with commercial/corporate interests – or that they have none to declare.
A.C. is a full time employee of Igenomix. A.D.M. was employee of Igenomix while working on this project. I.D.B., P.D.A., G.E., S.D.B. are full time employees of the MeriGen Research. All other authors declare that they have no conflicts of interest.