\section*{Figure legends}

%Each legend must be self contained, with all symbols and abbreviations used in the figure defined.

\textbf{Figure 1 - Features of mothers of the embryos.} \textbf{(A)} Median age of the mother at the event. There is no significant difference between first and recurrent miscarriages. \textbf{(B)} Gestational age at the time of the pregnancy termination. There is no significant difference between first and recurrent cases. \textbf{(C)} Folic acid intake. Range of values of menarche age \textbf{(D)} and Body Mass Index \textbf{(E)} in mothers of the embryos are not significantly different from a control set of mothers undergoing voluntary termination of pregnancy.


\textbf{Figure 2 - Outcome of the screening for aneuploidies in the embryos.} Forty-six embryos were screened by quantitative PCR to determine aneuploidies of chromosomes 13, 15, 16, 18, 21, 22, X, and Y, as well as to determine maternal contamination. DNA from embryos with no anuploidies in these chromosomes were further analyzed by comparative hybridization and shallow sequencing. Overall, we found aneuploidies in 56.6\% of the embryos, the most common being the trisomy of chromosome 22. In yellow the fraction of euploid embryos. 

\textbf{Figure 3 - Overview of the pipeline for prioritization of the genetic variants.} \textbf{(A)} \gp takes as input information about genomic variants found in cases and controls and outputs a subset of variants prioritized according to user-defined parameters. \gp currently analyzes coding regions and performs four filtering steps. \textbf{(B)} Filter I retains variants based on three criteria: overall impact on the gene product (moderate or high), allele frequency in control populations, and the combined property of being putatively damaging (quantified by the CADD score) and located in genes intolerant to loss of function (determined by the pLIscore). It is also possible to incorporate one or more user-defined lists of genes relevant to the trait under study. \textbf{(C)} Filter III determines the chance for genes to be selected in a control population based on criteria specified in Filter I. In practice, a number of control individuals are sampled a number of times and their genetic data filtered using Filter I to obtain a list of genes selected by chance.


\textbf{Figure 4 - Features of the filtering steps.} \textbf{(A)} Number of hits per gene after Filter II. The majority of genes have less than five hits and there is no significant correlation between number of hits and gene length (Spearman r$^2$=0.05 p-value=0.124). In the insert: the genes with $>$5 hits are enriched for private variants. \textbf(B) Frequency across 100 replicates of genes that pass Filter I in resampling from a control population. Most genes are retained $<$5\% of times (yellow) therefore are retained if found in the embryos. 1,531 genes are instead retained in $>$5\% of replicates (blue) and therefore discarded if found in the embryos, under the assumption that they can be filtered by chance in healthy controls. \textbf{(C)} Despite comparable read depth between private and non-private variants after Filter III to control for possible artifacts due to low coverage, we further filter to remove hits that are private and with read depth outside the range found in non-private ones.

\textbf{Figure 5 - Results of the prioritization pipeline. (A)} Number of variants per embryo stratified by impact. Overall 4.1\% of the prioritized variants are classified as having high impact on the gene products. \textbf{(B)} Results per embryo. On the y-axis we give the prioritized genes while the x-axis shows the number of mutations per gene. Colors indicate the allele count and the class of severity. \textbf{(C)} Selection of prioritized variants/genes shared by embryos.


\textbf{Figure 6 - Two-SNP haplotype in \textit{FMNL2} prioritized in five embryos. (A)} Alelle frequencies from the NCBI Alpha database of the alternate alleles at rs750755379 and rs866373641. The two alleles exist at moderate-to-high frequency in human populations. \textbf{(B)} Position in the protein of rs750755379 (stop gain) and rs866373641 (missense). The stop-gain mutation is located in the Rho GTPase-binding/formin homology 3 (GBD/FH3) domain involved in subcellular localization and regulation of activation. The resulting truncated protein lacks the Formin Homology-2 (FH2) domain, which directly binds to the actin filament catalyzing its nucleation and elongation.

\textbf{Supplementary Figure 1 - Optimization of tissue homogenization and DNA extraction.} We do not observe significant difference between two methods of tissue homogenization (\textbf{A}), and three methods of DNA isolation (\textbf{b}) apart form a slightly higher range of yield for one type of resin. VTP: voluntary pregnancy termination;  FPL: first pregnancy loss; RPL:recurrent pregnancy loss;  MAT: maternal bllod; PoC: product of conception.


\textbf{Supplementary Figure 2 - Principal Component Analysis.} Plot of the first and second components obtained using 1,2M autosomal SNPs and considering genomic data of the ten embryos sequenced in this study and publicly available data of individuals form the Human Genome Diversity Project